Mitigating Interest Flooding DDoS Attacks in NDN

### 2013-01-11 03-interest-flooding.tex
Algorithm renaming, pseudocodes, consistency in "Interest" word usage, and some minor typos and grammars
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@894 df1a90cd-247b-4364-bdaf-287a704eb499

Unlike data packets, interest packets in NDN are unsolicited, consume state in intervening routers and are routed based on content name prefixes in the network. Such properties makes interest packets an effective tool in launching DoS attacks in NDN. An attacker or a set of distributed attackers can inject excessive amounts of Interests to overload the network and cause service disruptions for legitimate users (Fig.~\ref{fig:flooding example}). We coin in the term \emph{Interest flooding} to refer to these network-level flooding attacks.

@@

To efficiently implement an Interest flooding attack targeting a namespace (e.g., /newtorktimes/), an adversary needs to make sure that (1) the expressed Interests are getting as close to the Data producer as possible, and (2) the corresponding PIT entries (state) are kept on NDN routers for as long as possible \footnote{While NDN allows users to specify lifetime of the Interest inside the Interest packet~\cite{ndn-conext,ndn-tr}, it is router's decision for how long it is willing to keep the PIT entry.  For example, and we assume this in the rest of the paper, the maximum time for which any Interest is admitted is one second.}

In order to achieve both requirements most effectively, malicious Interests should avoid collapsing (i.e, request a content that is different than existing PIT entries in the network) and cache hits (i.e., request unpopular or non-existing content). In other words, to maximize effect of an DDoS attack, each malicious Interest generated by an individual attack bot should request a non-existing content with a unique name (\emph{unique junk Interests}). In this paper, we exclusively focus on this particular attack strategy as it maximizes the damaging effect of each malicious interest in the network and the same or similar countermeasures apply to all less-effective strains of the attack.

In the rest of this paper, we use the general term \emph{interest flooding attack} to refer to the above described attack and assume an attacker that is limited to controlling a botnet of end-hosts only (i.e, we assume the routers in the network and the computers in the victim domain are not compromised).

>>>>

Unlike data packets, Interest packets in NDN are unsolicited, consume state in intervening routers and are routed based on content name prefixes in the network. Such properties makes Interest packets an effective tool in launching DoS attacks in NDN. An attacker or a set of distributed attackers can inject excessive amounts of Interests to overload the network and cause service disruptions for legitimate users (Fig.~\ref{fig:flooding example}). We coin in the term \emph{Interest flooding} to refer to these network-level flooding attacks.

@@

To efficiently implement an Interest flooding attack targeting a namespace (e.g., /newtorktimes/), an adversary needs to make sure that (1) the expressed Interests are getting as close to the Data producer as possible, and (2) the corresponding PIT entries (state) are kept on NDN routers for as long as possible.\footnote{While NDN allows users to specify lifetime of the Interest inside the Interest packet~\cite{ndn-conext,ndn-tr}, it is router's decision for how long it is willing to keep the PIT entry.  For example, and we assume this in the rest of the paper, the maximum time for which any Interest is admitted is one second.}

In order to achieve both requirements most effectively, malicious Interests should avoid collapsing (i.e, request a content that is different than existing PIT entries in the network) and cache hits (i.e., request unpopular or non-existing content). In other words, to maximize effect of an DDoS attack, each malicious Interest generated by an individual attack bot should request a non-existing content with a unique name (\emph{unique junk Interests}). In this paper, we exclusively focus on this particular attack strategy as it maximizes the damaging effect of each malicious Interest in the network and the same or similar countermeasures apply to all less-effective strains of the attack.

In the rest of this paper, we use the general term \emph{Interest flooding attack} to refer to the above described attack and assume an attacker that is limited to controlling a botnet of end-hosts only (i.e, we assume the routers in the network and the computers in the victim domain are not compromised).

### 2013-01-11 04a-design-naive.tex
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@893 df1a90cd-247b-4364-bdaf-287a704eb499
Algorithm renaming, pseudocodes, consistency in "Interest" word usage, and some minor typos and grammars
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@894 df1a90cd-247b-4364-bdaf-287a704eb499

In this section we present several algorithms to mitigate Interest flooding attacks in NDN.  Our mitigation strategies feature varying degrees of implementation complexity and effectiveness -- the higher the implementation complexity, the more effective is the algorithm against Interest flooding attacks. We start by describing our simple strategies and use the insights and lessons learnt from the deployment of these techniques to inform and design more effective mitigation techniques that work well in a variety of attack scenarios.

@@

An obvious and naive solution to defend against Interest flooding attacks is to restrict the number of Interests that are forwarded in the network. To this end, we exploit a 
 fundamental principle of NDN architecture - flow balance between Interest and Data packets. The flow balance refers to the fact that one Interest can be satisfied by at most one Data packet. This principle allows intermediate routers that forward the Interest packets to control the number of outstanding Interests in the network. 
One simple and easy method to implement this strategy is for a NDN router to limit the number of forwarded Interests out of each interface based on the physical capacity of the corresponding out interface. This technique is a slight modification of the well-known {\it Token Bucket} algorithm that is currently widely used in packet switched networks. Analogous to the {\it Token Bucket} algorithm, NDN routers keep track of the amount of Data requested that can fully utilize the downstream link (estimated from the number of Interests it forwarded) and once the link capacity limit has been reached, they no longer forward new incoming Interests. Ideally, the number of tokens {\it i.e.} the \emph{Pending Interests Limit} for each link will be proportional to the link's bandwidth-delay product (BDP)~\cite{tcp-survey}. We can formalize this value as follows:
% With the objective to request as many Data packets, as downstream link can pump through, we are getting the following equation for Interest limit:

@@

\subsubsection{\textbf{Physical limits with per-interface fairness}}
{\it Priya: Rename this algo to Token Bucket with per-interface fairness or something similar}
\label{sec:queuing}

>>>>

In this section we present several algorithms to mitigate Interest flooding attacks in NDN.  Our mitigation strategies feature varying degrees of implementation complexity and effectiveness---the higher the implementation complexity, the more effective is the algorithm against Interest flooding attacks. We start by describing our simple strategies and use the insights and lessons learnt from the deployment of these techniques to inform and design more effective mitigation techniques that can work well in a variety of attack scenarios.

@@

An obvious and na\"ive solution to defend against Interest flooding attacks is to restrict the number of Interests that are forwarded in the network. To this end, we exploit a 
 fundamental principle of NDN architecture---flow balance between Interest and Data packets. The flow balance refers to the fact that one Interest can be satisfied by at most one Data packet. This principle allows intermediate routers that forward the Interest packets to control the number of outstanding Interests in the network. 
One simple and easy method to implement this strategy is for a NDN router to limit the number of forwarded Interests out of each interface based on the physical capacity of the corresponding out interface. This technique is a slight modification of the well-known {\it Token Bucket} algorithm that is currently widely used in packet switched networks. Analogous to the {\it Token Bucket} algorithm, NDN routers keep track of the amount of Data requested that can fully utilize the downstream link (estimated from the number of Interests it forwarded) and once the link capacity limit has been reached, they no longer forward new incoming Interests. Ideally, the number of tokens {\it i.e.} the \emph{Pending Interests Limit} for each link will be proportional to the link's bandwidth-delay product (BDP)~\cite{tcp-survey}. We can formalize this value as follows:
% With the objective to request as many Data packets, as downstream link can pump through, we are getting the following equation for Interest limit:

@@

\subsubsection{\textbf{Token bucket with per interface fairness}}
\label{sec:queuing}

### 2013-01-11 04b-design-intelligent.tex
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@893 df1a90cd-247b-4364-bdaf-287a704eb499
Algorithm renaming, pseudocodes, consistency in "Interest" word usage, and some minor typos and grammars
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@894 df1a90cd-247b-4364-bdaf-287a704eb499

In order to distinguish between legitimate and malicious Interests, we leverage another unique feature of  NDN architecture --- guaranteed symmetric flow of Interest and Data packets. Since a Data packet takes the reverse path of the corresponding Interest packet, a router is guarateed to see if an Interest it forwarded resulted in a matching Data packet. The only exception being if Interest/Data packets were lost along the way due to congestion in the network.
Therefore, intermediate routers can classify all Interests that result in matching Data as legitimate, while the ones that timed out can be marked as malicious.\footnote{Recall that in order to maximize effect of the Interest flooding attack, an adversary expresses a large volume of junk Interests (see Section~\ref{sec:interest flooding}).}  
%Implications of other types of attacks are discussed in Section~\ref{sec:discussion}.}

This timeout-based differentiation method is reactive by nature: one cannot determine in advance if an Interest will result in a timeout or in Data being retrieved. However, routers can proactively maintain up-to-date statistics of Interests satisfaction ratios (number of forwarded versus number of satisfied Interests), and use this statistic to determine whether an incoming Interest must be forwarded or dropped. For example, maintaining independent Interest satisfaction ratio statistics for each incoming interface is sufficient to estimate whether an Interest received from a neighboring interface will result in a Data packet or a timeout. Statistics can also be kept at finer granularities such as per outgoing interface, per name prefix, etc. that can further improve the estimates. A router's goal is to penalize malicious Interests as soon as possible -- negative statistics should be allowed to build up fast, while positive statistics should not be allowed to deteriorate too quickly. Choosing the right balance between these contradictory requirements is a challenge and we explore this topic further  in the evaluation section.  {\it Priya: we should address the one that works best such as exponentially weighted moving average in eval section - it doesn't belong here.}

@@

\subsubsection{\textbf{Probabilistic Interest accept}}
\label{sec:probabilistic}

@@

\subsubsection{\textbf{Dynamic limits}}
\label{sec:dynamic limits}

>>>>

In order to distinguish between legitimate and malicious Interests, we leverage another unique feature of  NDN architecture---guaranteed symmetric flow of Interest and Data packets. Since a Data packet takes the reverse path of the corresponding Interest packet, a router is guarateed to see if an Interest it forwarded resulted in a matching Data packet. The only exception being if Interest/Data packets were lost along the way due to congestion in the network.
Therefore, intermediate routers can classify all Interests that result in matching Data as legitimate, while the ones that timed out can be marked as malicious.\footnote{Recall that in order to maximize effect of the Interest flooding attack, an adversary expresses a large volume of junk Interests (see Section~\ref{sec:interest flooding}).}  
%Implications of other types of attacks are discussed in Section~\ref{sec:discussion}.}

This timeout-based differentiation method is reactive by nature: one cannot determine in advance if an Interest will result in a timeout or Data being retrieved. However, routers can proactively maintain up-to-date statistics of Interests satisfaction ratios (number of forwarded versus number of satisfied Interests), and use this statistic to determine whether an incoming Interest must be forwarded or dropped. For example, maintaining independent Interest satisfaction ratio statistics for each incoming interface is sufficient to estimate whether an Interest received from a neighbor connected to this interface will result in a Data packet or a timeout. Statistics can also be kept at finer granularities such as per outgoing interface, per name prefix, etc. that can further improve the estimates. A router's goal is to penalize malicious Interests as soon as possible---negative statistics should be allowed to build up fast, while positive statistics should not be allowed to deteriorate too quickly. Choosing the right balance between these contradictory requirements is a challenge and we explore this topic further  in the evaluation section.  {\color{red} \it Priya: we should address the one that works best such as exponentially weighted moving average in eval section - it doesn't belong here.}

@@

\subsubsection{\textbf{Satisfaction-based Interest acceptance}}
\label{sec:probabilistic}

@@

\subsubsection{\textbf{Satisfaction-based pushback}}
\label{sec:dynamic limits}

### 2013-01-11 05-evaluation.tex
Most of the evaluation section is semi-done.  I suspect I put too much details and discussion in the evalutions... Not sure
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@871 df1a90cd-247b-4364-bdaf-287a704eb499
Algorithm renaming, pseudocodes, consistency in "Interest" word usage, and some minor typos and grammars
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@894 df1a90cd-247b-4364-bdaf-287a704eb499

% Metrics
Quality of the attack mitigation methods directly corresponds to the quality of service for the legitimate users during an ongoing attack, which in NDN network can be quantified through percentage of satisfied Interest.
For example, if the network implements a mitigation method $X$ and under the attack majority of user-expressed Interests are getting satisfied, then the method $X$ can be seen as highly effective.
At the same time, if only a small percentage of the expressed Interest are getting satisfied during the attack, the method $X$ can be called ineffective.

% Traffic pattern
In our experiments we made a simplistic assumption that both legitimate users and adversaries express Interests with constant average rates with randomized (uniform distribution) time gaps between each two consecutive Interests.
% Alex: not sure how to argue about the traffic pattern
Although this pattern is not truly realistic, it gives us a good statistical mix of traffic from all network users without excessive network buffering.
In addition to that, in each run of the simulation we ensure that all Interests expressed by the legitimate users during a no-attack period are satisfied.
For simplicity and without losing any generality, we also equalized the average rates with which the legitimate users express Interests.

>>>>

% Metrics
% not sure if i already explained that. our essential qualitative metric is .  
% To quantify the effectiveness of the mitigation mechanism this  metric we use satisfaction percentages of user Interests.
Quality of the attack mitigation methods directly corresponds to the quality of service for the legitimate users during an ongoing attack, which in NDN network can be quantified through percentage of satisfied Interest.
For example, if the network implements a mitigation method $X$ and under the attack majority of user-expressed Interests are getting satisfied, then the method $X$ can be seen as highly effective.
At the same time, if only a small percentage of the expressed Interest are getting satisfied during the attack, the method $X$ can be called ineffective.

% Traffic pattern
In our experiments we made a simplistic assumption that both legitimate users and adversaries express Interests with constant average rates with randomized (uniform distribution) time gaps between each two consecutive Interests.
% Alex: not sure how to argue about the traffic pattern
Although this pattern may seem not truly realistic, it approximates a good statistical mix of traffic from all network users without excessive network buffering.

In addition to that, in each run of the simulation we ensure that all Interests expressed by the legitimate users during a no-attack period are satisfied.
For simplicity, we also equalized the average rates with which the legitimate users express Interests.
At the same time, the attackers are sending junk Interests as much as they can get through (remember, that Interest limits will not allow real flooding of Intersests in all of the designed mitigation algorithms).

### 2013-01-11 algos/dynamic-limits.tex
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@893 df1a90cd-247b-4364-bdaf-287a704eb499
Algorithm renaming, pseudocodes, consistency in "Interest" word usage, and some minor typos and grammars
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@894 df1a90cd-247b-4364-bdaf-287a704eb499

Essentially, the probabilistic Interest accept algorithm divides the available forwarding tokens between interfaces in proportion to their Interest satisfaction ratios.
Given the drawbacks of this algorithm, the same effect can be achieved without using probabilistic methods. An alternate strategy is to enable and enforce additional Interest limits for each incoming interface, where the values of the limits directly depend on the interface satisfaction ratios.
Additionally, routers may need to explicitly announce these limits to their downstream neighbors, ensuring that any forwarded Interest from the downstream is actually getting through, resulting in genuine Interest satisfaction statistics.

The formal definition of the dynamic limits algorithm is presented in Pseudocode~\ref{alg:dynamic limits}, while Fig.~\ref{fig:dynamic limits example} illustrates how the algorithm would work in our toy example on Fig.~\ref{fig:flooding example}.
Assuming the initial (physical) limit $L=10$ and the current satisfaction ratios on the the router A are 50\% for \texttt{eth1} and 0\% for \texttt{eth0}, and on the router B the ratio is 30\%  \texttt{eth0}, each node would set and announce the following  incoming interface limits $L'$: 
\begin{enumerate}
\item the router B would set and announce the incoming interface limit $L'=3$;
\item the router A, after receiving announcement from B would readjust its incoming interface limits to $L'_{eth1} = 1.5$ and $L'_{eth0} = 0$; and
\item legitimate user and adversary may either obey or ignore the announced limit, which will be in any case enforced by the router A.
\end{enumerate}

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.3]{dynamic-limits}
  \caption{Dynamic limits example: routers explicitly tell neighbors how many Interest can be for sure delivered to the Data producer}
  \label{fig:dynamic limits example}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{algorithm}[h]
\caption{Dynamic limits}
\label{alg:dynamic limits}
\begin{algorithmic}[1]
\State{} \Comment{Same initialization, InData and Timeout functions as in Physical Limits algorithm}
\vspace{0.2cm}
\State{$\forall f \in \mathrm{interfaces} : L'_{f} \leftarrow L_{f}$} \Comment{Per-incoming interface Interest limit}

\vspace{0.2cm}

\State{} \Comment{\textit{Announcement from the neighbor}}
\Function{InLimits}{InInterface $if$, Limit $L'$}
    \State $L_{if} \leftarrow L'$
\EndFunction

\vspace{0.2cm}

\Function{AnnounceLimits}{} \Comment{\textit{E.g., every second}}
\For{\textbf{each} outgoing interface $of$}

   \For{\textbf{each} incoming interface $if$}
        \State $L'_{if}= {L_{of}} \times (1 - U_{if}/F_{if})$
        \State AnnounceLimit($if$, $L'_{if}$)
   \EndFor

@@

The zero limit for the adversary's link means that the router A is temporarily not willing to accept any interests from this link, not until the statistics decays to the appropriate level (recall Fig.~\ref{fig:ratio example}).
At the next iterations of the dynamic limits algorithm, the legitimate user will be able to gradually improve the statistics on both the router A and router B (all his Interests will get through and will return Data), eventually resulting in a full allowance ($L'=L=10$) in the links between the routers A and B, and the user and the router A.

It should be noted that while in the description of the dynamic limits algorithm we used notions of ``outgoing'' and ``incoming'' interfaces, in the real system all interfaces can be both incoming and outgoing.
Thus, it may not be entirely clear which outgoing limit $L_{of}$ (line 10 in the algorithm) should be used to calculate the incoming limit $L_{if}$.
To overcome this problem, in our actual implementation we enforced separate incoming/outgoing interface limits for each individual FIB entry.
That is, for each FIB entry we set a separate Interest limit for each incoming interface ($L_{if}^{'FIB}$) based on a sum of FIB entry limits for each outgoing interface $L=\sum{L_{of}^{FIB}}$.

% Alex: not sure if this paragraph belongs here
Essentially, both probabilistic Interest accept and the dynamic limits algorithm are forms of a well-known push-back mechanism~\cite{Pushback}, with several core differences.
First, we are suppressing (pushing back) unwanted requests for Data, not the actual Data.
Second, differentiating between good and bad Interests is based on the traffic symmetry property of NDN.
% Alex: I'm not entirely sure about this point... 
Finally, both intelligent attack mitigation algorithms can be enabled all the time, without degrading network performance when there are no active attack.

>>>>

Essentially, the satisfaction-based Interest acceptance algorithm divides the available forwarding tokens between interfaces in proportion to their Interest satisfaction ratios.
% Given the drawbacks of this algorithm, the same effect can be achieved without relying on probabilistic models. 
An alternate strategy, which does not have described drawbacks, is to enable and enforce explicit Interest limits for each incoming interface, where the values of the limits directly depend on the interface satisfaction ratios.
Routers need to announce these limits to their downstream neighbors, ensuring that any forwarded Interest from the downstream is actually getting through, resulting in genuine Interest satisfaction statistics.

The formal definition of the satisfaction-based pushback algorithm is presented in Pseudocode~\ref{alg:dynamic limits}, while Fig.~\ref{fig:dynamic limits example} illustrates how the algorithm would work in our toy example on Fig.~\ref{fig:flooding example}.
Assuming the initial token bucket limit $L=10$ and the current satisfaction ratios on the the router A are 50\% for \texttt{eth1} and 0\% for \texttt{eth0}, and on the router B the ratio is 30\% for \texttt{eth0}, each node would set and announce the following  incoming interface limits $L'$: 
\begin{enumerate}
\item the router B would set and announce the incoming interface limit $L'=3$;
\item the router A, after receiving announcement from B would readjust its incoming interface limits to $L'_{eth1} = 1.5$ and $L'_{eth0} = 0$; and
\item legitimate user and adversary may either obey or ignore the announced limit, which will be in any case enforced by the router A.
\end{enumerate}

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.3]{dynamic-limits}
  \caption{Satisfaction-based pushback example: routers explicitly tell neighbors how many Interest can be for sure delivered to the Data producer}
  \label{fig:dynamic limits example}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{ 
\begin{algorithm}[h]
\footnotesize
\caption{\small Satisfaction-based pushback}
\label{alg:dynamic limits}
\begin{algorithmic}[1]
% \State{}\Comment{Same initialization, InData and Timeout functions as in Pseudocode~\ref{alg:queuing}}
\State{} \Comment{Same init, InData and Timeout functions as in Pseudocode~\ref{alg:queuing}}
\vspace{0.1cm}
\State{$\forall f \in \mathrm{interfaces} : L'_{f} \leftarrow L_{f}$} \Comment{Per-incoming interface Interest limit}

\vspace{0.1cm}

\State{} \Comment{\textit{Announcement from the neighbor}}
\Function{InLimits}{InInterface $in$, Limit $L'$}
    \State $L_{in} \leftarrow L'$
\EndFunction

\vspace{0.1cm}

\Function{AnnounceLimits}{} \Comment{\textit{E.g., every second}}
\For{\textbf{each} outgoing interface $out$}

   \For{\textbf{each} incoming interface $in$}
        \State $L'_{in}= {L_{out}} \times (1 - U_{in}/F_{in})$
        \State AnnounceLimit($in$, $L'_{in}$)
   \EndFor

@@

The zero limit for the adversary's link means that the router A is temporarily not willing to accept any Interests from this link, not until the statistics decays to the appropriate level (recall Fig.~\ref{fig:ratio example}).
At the next iterations of the satisfaction-based pushback algorithm, the legitimate user will be able to gradually improve the statistics on both the router A and router B (all his Interests will get through and will return Data), eventually resulting in a full allowance ($L'=L=10$) in the links between the routers A and B, and the user and the router A.

It should be noted that while in the description of the satisfaction-based pushback algorithm we used notions of ``outgoing'' and ``incoming'' interfaces, in the real system all interfaces can be both incoming and outgoing.
Thus, it may not be entirely clear which outgoing limit $L_{out}$ (line 10 in the algorithm) should be used to calculate the incoming limit $L_{in}$.
To overcome this problem, in our actual implementation we enforced separate incoming/outgoing interface limits for each individual FIB entry.
That is, for each FIB entry we set a separate Interest limit for each incoming interface ($L_{in}^{'FIB}$) based on a sum of FIB entry limits for each outgoing interface $L=\sum{L_{out}^{FIB}}$.

% Alex: not sure if this paragraph belongs here
Essentially, both satisfaction-based algorithm are forms of a well-known push-back mechanism~\cite{Pushback}, with several core differences.
First, we are suppressing (pushing back) unwanted requests for Data, not the actual Data.
Second, differentiating between good and bad Interests is based on the traffic symmetry property of NDN.
% Alex: I'm not entirely sure about this point... 
Finally, both intelligent attack mitigation algorithms can be enabled all the time, without degrading network performance when there is no active attack.

### 2013-01-11 algos/fair-queueing.tex
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@893 df1a90cd-247b-4364-bdaf-287a704eb499
Algorithm renaming, pseudocodes, consistency in "Interest" word usage, and some minor typos and grammars
git-svn-id: https://ndnpub.parc.com/ndn-papers/NDN-SEC/dos-protection@894 df1a90cd-247b-4364-bdaf-287a704eb499

To address the lack of fairness associated with the naive {\it Token Bucket} algorithm, we modify it to ensure that the Interests forwarded by a router on each interface represent a fair mix of Interests received from neighboring interfaces. For example, in Fig.~\ref{fig:flooding example} router A can ensure that the tokens associated with Interests sent out on interface {\texttt eth2}  are fairly distributed across incoming interfaces \texttt{eth0} and \texttt{eth1}. 
%Due to the expected small volume of Interests, we cannot merely rely on network buffers to perform statistical multiplexing of Interests, as they would almost never be %buffered.
%At the same time, until bag of tokens is not empty, there is no reason to delay Interest forwarding, as we do not known how many and from which interfaces Interests %will arrive in the future.
In order to achieve our goal of ensuring ``fair'' mixing of Interests from all neighboring interfaces, 
%, we implement an additional features to support buffering of Interests, if they cannot be immediately forwarded. 
we extend the Pending Interest Table to support flagging of Interests that cannot be immediately forwarded (see Fig.~\ref{fig:queueing}). To support fair queuing, we implement hierarchical queues for each interface as shown in Fig.~\ref{fig:queueing})\footnote{This mechanism is essentially class based queuing, with classes for each outgoing/incoming interface.} As an alternative to hierarchical queues, one can also use an approach based on virtual time. For details, please see ~\cite{zhang1990virtual}.  We note that unlike normal queuing, Interest queues do not actually store a packet, but merely a bi-directional pointer to the existing PIT entry.
Thus, a PIT entry can be quickly updated when the Interest is actually forwarded; further the element can be quickly removed from the queue when the Interest expires.

@@

We present a more formalized description of this algorithm in Pseudocode~\ref{alg:queuing}. We can control the amount of physical resources utilized at a router (such as memory and computation power) by setting appropriate queue sizes. It is also important to set apprpriate value for how long an Interest can be enqueued. If an Interest is enqueued for a long time, by the time it is dequeued and appropriately forwarded, the resulting Data packet might not be valid due to state expiration at previous downstream routers. We believe that implementing additional mechanisms on each pair of communicating routers to keep the Interests from getting stale would mitigate this issue, though the details are beyond the scope of this paper.
{\it Priya:  We need to  touch upon the Interest freshness issue in the NDN overview section; else we should get rid of this bit.}

@@

\begin{algorithm}[h]
\caption{Physical limits with per-interface fairness}
\label{alg:queuing}
\begin{algorithmic}[1]
\State{} \Comment{Same initialization, InData and Timeout functions as in Physical Limits algorithm}

\vspace{0.2cm}

\Function{OutInterest}{Interest \textbf{i}, InInterface \textbf{if}, OutInterface \textbf{of}}
    \If{$L_{of} - O_{of} > 0$} \Comment{\textbf{of} is under physical limits}
        \State $O_{of} \leftarrow O_{of} + 1$  \Comment{``Borrow'' token}
        \State add \textbf{of} to PIT entry and forward \textbf{i} to \textbf{of}
    \Else
        \State Queue $q \leftarrow of$.GetSubQueue($if$)
        \If{$Size(q) < L_{of}$}
           \State $q$.PushInterest($i$)
           \State add \textbf{of} to PIT entry, and link PIT entry with the queue
        \Else
           \State drop Interest
        \EndIf
    \EndIf
\EndFunction

\vspace{0.2cm}
\State{} \Comment{\textit{Whenever $L_{of} - O_{of}$ becomes larger than zero}}
\Function{TokenBecomesAvailable}{}
    \State Queue $q \leftarrow$ $of$.GetRoundRobinSubQueue 
    \State $i \leftarrow$ $q$.PopInterest
    \State update PIT entry and Forward($i$, $of$)
\EndFunction
\end{algorithmic}
\end{algorithm}

As we show in Section~\ref{sec:evaluation}, fair queueing provides partial relief from the Interest flooding attack, allowing legitimate users to successfully fetch Data for 15--20\% of the expressed Interests (compared to 0-10\% without fair queueing). While this algorithm might be reasonable for ensuring fairness in an NDN network, it is completely ineffective in protecting legitimate users from malicious attackers. Malicious attackers are able to successfully thwart access to content for legitimate users by sending a relatively modest volume of Interests.

>>>>

To address the lack of fairness associated with the na\"ive {\it Token Bucket} algorithm, we modify it to ensure that the Interests forwarded by a router on each interface represent a fair mix of Interests received from neighboring interfaces. For example, in Fig.~\ref{fig:flooding example} router A can ensure that the tokens associated with Interests sent out on interface {\texttt eth2}  are fairly distributed across incoming interfaces \texttt{eth0} and \texttt{eth1}. 
%Due to the expected small volume of Interests, we cannot merely rely on network buffers to perform statistical multiplexing of Interests, as they would almost never be %buffered.
%At the same time, until bag of tokens is not empty, there is no reason to delay Interest forwarding, as we do not known how many and from which interfaces Interests %will arrive in the future.
In order to achieve our goal of ensuring ``fair'' mixing of Interests from all neighboring interfaces, 
%, we implement an additional features to support buffering of Interests, if they cannot be immediately forwarded. 
we extend the Pending Interest Table to support flagging of Interests that cannot be immediately forwarded (see Fig.~\ref{fig:queueing}). To support fair queuing, we implement hierarchical queues for each interface as shown in Fig.~\ref{fig:queueing})\footnote{This mechanism is essentially class based queuing, with classes for each outgoing/incoming interface.} As an alternative to hierarchical queues, one can also use an approach based on virtual time. For details, please see ~\cite{zhang1990virtual}.  We note that unlike normal queuing, Interest queues do not actually store a packet, but merely a bi-directional pointer to the existing PIT entry.
Thus, a PIT entry can be quickly updated when the Interest is actually forwarded; further the element can be quickly removed from the queue when the Interest expires.

@@

We present a more formalized description of this algorithm in Pseudocode~\ref{alg:queuing}. We can control the amount of physical resources utilized at a router (such as memory and computation power) by setting appropriate queue sizes. It is also important to set apprpriate value for how long an Interest can be enqueued. If an Interest is enqueued for a long time, by the time it is dequeued and appropriately forwarded, the resulting Data packet might not be valid due to state expiration at previous downstream routers. We believe that implementing additional mechanisms on each pair of communicating routers to keep the Interests from getting stale would mitigate this issue, though the details are beyond the scope of this paper.
{\color{red} \it Priya:  We need to  touch upon the Interest freshness issue in the NDN overview section; else we should get rid of this bit.}

@@

\begin{algorithm}[h]
\footnotesize
\caption{\small Token bucket with per interface fairness}
\label{alg:queuing}
\begin{algorithmic}[1]
\For{\textbf{each} interface \textbf{if}}
    \State{$L_{if} \leftarrow$ Interest Limit according to (1)}
    \State{$O_{if} \leftarrow 0$} \Comment{Outstanding Interests on interface \textbf{if}}
\EndFor

\vspace{0.1cm}

\Function{OutInterest}{Interest \textbf{i}, InInterface \textbf{in}, OutInterface \textbf{out}}
    \If{$L_{out} - O_{out} > 0$} \Comment{\textbf{out} is under limit cap}
        \State $O_{out} \leftarrow O_{out} + 1$  \Comment{``borrow'' a token from the bucket}
        \State add \textbf{out} to PIT entry and forward \textbf{i} to \textbf{out}
    \Else
        \State Queue $q \leftarrow out$.GetSubQueue($in$)
        \If{$Size(q) < L_{out}$}
           \State $q$.PushInterest($i$)
           \State add \textbf{out} to PIT entry, and link PIT entry with the queue
        \Else
           \State drop Interest
        \EndIf
    \EndIf
\EndFunction

\vspace{0.1cm}

\State{} \Comment{\textit{Whenever $L_{out} - O_{out}$ becomes larger than zero}}
\Function{TokenBecomesAvailable}{}
    \State Queue $q \leftarrow$ $out$.GetRoundRobinSubQueue 
    \State Interest $i \leftarrow$ $q$.PopInterest
    \State update PIT entry and Forward($i$, $out$)
\EndFunction

\vspace{0.1cm}

\Function{InData}{Data \textbf{d}}
   \State lookup PIT entry \textbf{p} for data \textbf{d}
   \For{\textbf{each} outgoing interface \textbf{out} in \textbf{p}}
        \State $O_{out} \leftarrow O_{out} - 1$ \Comment{``return'' token}
   \EndFor
\EndFunction

\vspace{0.1cm}

\Function{Timeout}{PIT entry $p$}
   \For{\textbf{each} outgoing interface \textbf{out} in \textbf{p}}
        \State $O_{out} \leftarrow O_{out} - 1$ \Comment{``return'' token}
   \EndFor
\EndFunction

\end{algorithmic}
\end{algorithm}

As we show in Section~\ref{sec:evaluation}, fair queueing provides partial relief from the Interest flooding attack, allowing legitimate users to successfully fetch Data for 15--20\% of the expressed Interests (compared to 0--10\% without fair queueing). While this algorithm might be reasonable for ensuring fairness in an NDN network, it is completely ineffective in protecting legitimate users from malicious attackers. Malicious attackers are able to successfully thwart access to content for legitimate users by sending a relatively modest volume of Interests.
